SVM, K-NN

### 2021-02-23 SVM,kNN.Rmd

---
title: "Multivariate Analysis"
author: "Scott Coffin"
date: "1/19/2021"
output: 
  html_document:
    code_folding: hide
    theme: sandstone
    toc: yes
    toc_float: yes
    toc_depth: 4
    number_sections: true
  word_document:
    toc: yes
---
```{r global-options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      warning=FALSE, message=FALSE, echo = FALSE)
```
---

# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load packages
library(tidyverse) #General everything
library(RColorBrewer)
library(ggplot2) #General plotting
library(ggrepel) #For adding text labels that repel away from data points
library(calecopal) #Color palette
library(shiny) #Runs shiny
library(shinythemes) #Shiny theme for the page
library(shinyWidgets) #Widgets
library(scales) #SSD - Use the percent format
library(reshape2) #Overview tab - melts bars together
library(ssdtools) #SSD package
library(DT) #Build HTML data tables
#library(plotly) #Make plots interactive
library(viridis) #Colors
library(scales) #To use "percent" function
library(tigerstats) #turns things into percents
library(ggbeeswarm) #plot all points
library(fitdistrplus, exclude = 'select') #alt SSD 
library(RColorBrewer) #colors
library(pheatmap) #pretty heat maps
library(rpart)  #for trees
#library(rattle)    # Fancy tree plot This is a difficult library to install (https://gist.github.com/zhiyzuo/a489ffdcc5da87f28f8589a55aa206dd) 
library(rpart.plot)             # Enhanced tree plots
library(RColorBrewer)       # Color selection for fancy tree plot
library(party)                  # Alternative decision tree algorithm
# library(partykit)               # Convert rpart object to BinaryTree
library(pROC)   #for ROC curves
library(uwot) #umap
library(ISLR)  #for the Carseat Data
library(lme4) #general linearized mixed model GLMM
library(quantregForest)
library(caret)
library(tidyverse)
library(tidymodels)
library(skimr)
library(sf)
library(ggspatial)
library(nhdplusTools)
library(patchwork)
library(Metrics)
library(gt)
library(randomForest)
library(DALEX) #explanatory model analysis

```
## Data import
```{r data import}
# Load finalized dataset.
aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000) #%>% 
  #mutate(ID = paste0("ID",row_number()))

@@

multiVar <- aoc_z %>% dplyr::select(#doi, size.category, 
                                    size_f,
                                    size.length.um.used.for.conversions, 
                                    shape, 
                                    polymer, 
                                    particle.volume.um3, 
                                    density.mg.um.3, 
                                    organism.group,
                                    environment, 
                                    bio.org, #biological level of organization
                                    #af.time, #assessment factor based on exposure time
                                    treatments, #number of doses (no including control)
                                    effect, #yes no
                                    effect_f,
                                    effect_10,
                                    size_f,
                                    exposure.duration.d, 
                                    exposure.route, #Factor
                                    organism.group, #factor
                                    media.temp, #numeric
                                    lvl1_f, #endpoints
                                    lvl2_f, #endpoints
                                   # lvl3, 
                                     dose.mg.L.master, 
                                    sex, #factor
                                    media.ph, #numeric
                                    media.sal.ppt, #numeric
                                    dose.particles.mL.master,
                                   effect.metric, #NOEC LOEC
                                    functional.group, #factor
                                    charge, #positive or negatibe
                                    zetapotential.mV, # numeric   
                                   max.size.ingest.mm,#max ingestible size
                                   acute.chronic_f,
                                   dose.mg.L.master.AF.noec,
                                   dose.particles.mL.master.AF.noec,
                                   dose.um3.mL.master,
                                   max.size.ingest.mm, #maximum ingestible size range
                                   effect.score) %>%  #1 = minor, 2 = photosynthesis, feeding, 3 = growth, chlorophyll content, 4 = reproduction, 5  = population growth, 6 = survival
  filter(!size_f == "Not Reported") %>%   #take out not reported 
                                 #  max.size.ingest.mm) %>%  #max ingestible size
  filter(!size_f == "Not Reported")  #take out not reported

@@

# Analysis
## Data Exploration
### Completeness vby Size
```{r completeness heatmap 1}
CompletenessSummary_size <- multiVar %>%
  group_by(size_f) %>% 
     summarise_all((name = ~sum(is.na(.))/length(.))) %>% 
  mutate(across(is.numeric,~round(., 2))) %>% 
  mutate(across(is.numeric, ~100 *(1 -.)))
CompletenessSummary_size
```

```{r}
require(pheatmap)
#convert to matrix and transpose
transposed_size <- as.data.frame(t(as.matrix(CompletenessSummary_size[2:25]))) %>% #1 is category, 2-6 are variables
  arrange('1nm < 100nm')
#reassign column names
colnames(transposed_size) <- c("1nm < 100nm", "100nm < 1µm", "1µm < 100µm", "100µm < 1mm", "1mm < 5mm")
#transposedTag
#format as matrix
MissingMatrix_size <- data.matrix(transposed_size)
#build heatmap
pheatmap(MissingMatrix_size,
                           main = "Data Completeness by Size Bin", #title
                           fontsize = 12,
                           cluster_rows = FALSE, cluster_cols = FALSE,#disable dendrograms
                           display_numbers = TRUE,
                           treeheight_row = 0, treeheight_col = 0, #keeps clustering after dropping dendrograms
                           col = rev(brewer.pal(n = 9, name = "PuBu"))) #blue color scheme with 9 colors)

```
### Completeness by Environment
```{r completeness heatmap 2}
CompletenessSummary_environment <- multiVar %>%
  filter(!environment == "NA") %>% 
  group_by(environment) %>% 
     summarise_all((name = ~sum(is.na(.))/length(.))) %>% 
  mutate(across(is.numeric,~round(., 2))) %>% 
  mutate(across(is.numeric, ~100 *(1 -.)))
CompletenessSummary_environment
```

```{r}
require(pheatmap)
#convert to matrix and transpose
transposed_environment <- as.data.frame(t(as.matrix(CompletenessSummary_environment[2:25])))
#reassign column names
colnames(transposed_environment) <- c("Freshwater", "Marine", "Terrestrial")
#transposedTag
#format as matrix
MissingMatrix_environment <- data.matrix(transposed_environment)
#build heatmap
pheatmap(MissingMatrix_environment,
                           main = "Data Completeness by environment", #title
                           fontsize = 12,
                           cluster_rows = FALSE, cluster_cols = FALSE,#disable dendrograms
                           display_numbers = TRUE,
                           treeheight_row = 0, treeheight_col = 0, #keeps clustering after dropping dendrograms
                           col = rev(brewer.pal(n = 9, name = "PuBu"))) #blue color scheme with 9 colors)

## Stepwise Regression

```{r}
## Estimate an OLS Regression
fitols <- lm(effect_10 ~ size.length.um.used.for.conversions + particle.volume.um3 + exposure.duration.d + media.temp + dose.particles.mL.master, 
             na.action = na.omit, 
             data = multiVar)
summary(fitols)
```

```{r}
require(reshape2)
multiVar %>% 
  dplyr::select(size.length.um.used.for.conversions, particle.volume.um3, exposure.duration.d ,media.temp, dose.particles.mL.master) %>% 
  melt() %>%  #convert wide to long
   mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
  ggplot(aes(x = value)) + 
  stat_density() + 
  facet_wrap(~variable, scales = "free")
```

<!-- # ```{r} -->
<!-- # # cdplot(as.factor(effect) ~ particle.volume.um3, data = multiVar, -->
<!-- #        main = "Conditional Density Plot of ", xlab = "Dose", ylab = "Effect") -->
<!-- # ``` -->

## Random Forest
### Recursive Partitioning And Regression Trees

The rpart algorithm works by splitting the dataset recursively, which means that the subsets that arise from a split are further split until a predetermined termination criterion is reached.  At each step, the split is made based on the independent variable that results in the largest possible reduction in heterogeneity of the dependent (predicted) variable.

It is important to note that the algorithm works by making the best possible choice at each particular stage, without any consideration of whether those choices remain optimal in future stages. That is, the algorithm makes a locally optimal decision at each stage. It is thus quite possible that such a choice at one stage turns out to be sub-optimal in the overall scheme of things.  In other words,  the algorithm does not find a globally optimal tree.
```{r}
#trim data so effect is always known
multiVar_sub <- multiVar %>% 
  drop_na(effect_10)

# Split data into training and test sets
set.seed(42)
multiVar_sub[,"train"] <- ifelse(runif(nrow(multiVar_sub)) < 0.8, 1, 0)
# Separate trainig and test sets
trainSet <- multiVar_sub[multiVar_sub$train==1,]
testSet <- multiVar_sub[multiVar_sub$train==0,]
#get column index of train flag
trainColNum <- grep("train", names(trainSet))
# Remove train flag column from train and test sets
trainSet <- trainSet[, -trainColNum]
testSet <- testSet[, -trainColNum]
```
Make a classification tree.

```{r, echo = FALSE}
#get column index of predicted variable in dataset
typeColNum <- grep("effect",names(multiVar_sub))

#build tree
require(rpart)
set.seed(15097)
t1 <- rpart(effect ~ size.length.um.used.for.conversions + particle.volume.um3 + exposure.duration.d + media.temp + dose.particles.mL.master,
            method = "class", #classification because response is discrete
            control = rpart.control(minbucket = 20, cp=0.008), #requires that there be at least 19 cases (responded + nonrespondents) in the final grouping of variable values of the terminal node of the tre..
            data = trainSet)
print(t1, digits=4)
```

Plot an interpretable tree.
```{r, echo = FALSE}
require(rpart.plot)
cols <- ifelse(t1$frame$yval == 1, "gray50", "black")
prp(t1, main="Tree for Effect",
    extra=106, # display prob of survival and percent of obs
    nn=TRUE, # display node numbers
    fallen.leaves=TRUE, # put leaves on the bottom of page
    branch=.5, # change angle of branch lines
    faclen=0, # do not abbreviate factor levels
    trace=1, # print automatically calculated cex
    shadow.col="gray", # shadows under the leaves
    branch.lty=1, # draw branches using solid lines
    branch.type=5, # branch lines width = weight(frame$wt), no. of cases here
    split.cex=1.2, # make split text larger than node text
    split.prefix="is ", # put "is " before split text
    split.suffix="?", # put "?" after split text
    col=cols, border.col=cols, # cols[2] if survived
    split.box.col="lightgray", # lightgray split boxes (default is white)
    split.border.col="darkgray", # darkgray border on split boxes
    split.round=0.5) # round the split box corners a tad
```
Next we see how good the model is by seeing how it fares against the test data.
```{r}
t1_predict <- predict(t1, newdata = testSet[,-typeColNum],
                      type="class")
mean(t1_predict==testSet$effect)
# [1] 0.7094088
#confusion matrix
table(pred=t1_predict,true=testSet$effect)
```
```{r}
par(mfrow=c(1,2)) # two plots on one page
#plot approximate R-squared and relative error for different splits (2 plots). labels are only appropriate for the "anova" method.
rsq.rpart(t1)
```

Next, we prune the tree using the cost complexity criterion. Basically, the intent is to see if a shallower subtree can give us comparable results. If so, we’d be better of choosing the shallower tree because it reduces the likelihood of overfitting.

As described earlier, we choose the appropriate pruning parameter (aka cost-complexity parameter) \alpha by picking the value that results in the lowest prediction error. Note that all relevant computations have already been carried out by R when we built the original tree (the call to rpart in the code above). All that remains now is to pick the value of \alpha:

#### Pruning

```{r}
#cost-complexity pruning
printcp(t1)
```
It is clear from the above, that the lowest cross-validation error (xerror in the table) occurs for \alpha =0.008 (this is CP in the table above).   One can find CP programatically like so:
```{r}
# get index of CP with lowest xerror
opt <- which.min(t1$cptable[,"xerror"])
#get its value
cp <- t1$cptable[opt, "CP"]
```
Next, we prune the tree based on this value of CP:

```{r}
# # prune the tree
# pt1 <- prune(t1,cp)
# pt1<- prune(t1, cp= t1$cptable[which.min(t1$cptable[,"xerror"]),"CP"])
# 
# # plot the pruned tree
# plot(pt1, uniform=TRUE,
#    main="Pruned Classification Tree");text(pt1, use.n=TRUE, all=TRUE, cex=.8)
# 
# #post(pfit, file = "c:/ptree.ps",
#  #  title = "Pruned Classification Tree for Kyphosis")
```
```{r}
# #find proportion of correct predictions using test set
# t1_pruned_predict <- predict(pt1, testSet, type="class")
# mean(t1_pruned_predict == testSet$effect)
# #
```
This is not an improvement over an unprunend tree.. We need to check that this holds up for different training and test sets. This is easily done by creating multiple random partitions of the dataset and checking the efficacy of pruning for each. To do this efficiently, I’ll create a function that takes the training fraction, number of runs (partitions) and the name of the dataset as inputs and outputs the proportion of correct predictions for each run. It also optionally prunes the tree.

```{r}
# <!-- # #function to do multiple runs -->
# <!-- # multiple_runs_classification <- function(train_fraction,n,dataset,prune_tree=FALSE){ -->
# <!-- # fraction_correct <- rep(NA,n) -->
# <!-- # set.seed(42) -->
# <!-- # for (i in 1:n){ -->
# <!-- #   dataset[,"train"] <- ifelse(runif(nrow(dataset))<0.8,1,0) -->
# <!-- #   trainColNum <- grep("train",names(dataset)) -->
# <!-- #   typeColNum <- grep("effect",names(dataset)) -->
# <!-- #   trainSet <- dataset[dataset$train==1,-trainColNum] -->
# <!-- #   testSet <- dataset[dataset$train==0,-trainColNum] -->
# <!-- #   rpart_model <- rpart(effect~ size.length.um.used.for.conversions + particle.volume.um3 + exposure.duration.d + media.temp + dose.particles.mL.master,data = trainSet, method="class") -->
# <!-- # if(prune_tree==FALSE) { -->
# <!-- #   rpart_test_predict <- predict(rpart_model,testSet[,-typeColNum],type="class") -->
# <!-- #   fraction_correct[i] <- mean(rpart_test_predict==testSet$effect) -->
# <!-- #   }else{ -->
# <!-- #     opt <- which.min(rpart_model$cptable[,"xerror"]) -->
# <!-- #     cp <- rpart_model$cptable[opt, "CP"] -->
# <!-- #     pruned_model <- prune(rpart_model,cp) -->
# <!-- #     rpart_pruned_predict <- predict(pruned_model,testSet[,-typeColNum],type="class") -->
# <!-- #     fraction_correct[i] <- mean(rpart_pruned_predict == testSet$effect) -->
# <!-- #   } -->
# <!-- #   } -->
# <!-- # return(fraction_correct) -->
# <!-- # } -->
# <!-- # ``` -->
#  
# <!-- # Note that in the above,  I have set the default value of the prune_tree to FALSE, so the function will execute the first branch of the if statement unless the default is overridden. -->
# <!-- #  -->
# <!-- # OK, so let’s do 50 runs with and without pruning, and check the mean and variance of the results for both sets of runs. -->
# <!-- # ```{r} -->
# <!-- # #50 runs, no pruning -->
# <!-- # unpruned_set <- multiple_runs_classification(0.8,50,multiVar_sub) -->
# <!-- # mean(unpruned_set) -->
# <!-- # #[1] 0.7261882 -->
# <!-- # sd(unpruned_set) -->
# <!-- # #[1] 0.01554347 -->
# <!-- # #50 runs, with pruning -->
# <!-- # pruned_set <- multiple_runs_classification(0.8,50,multiVar_sub,prune_tree=TRUE) -->
# <!-- # mean(pruned_set) -->
# <!-- # #[1] 0.7261875 -->
# <!-- # sd(pruned_set) -->
# <!-- # #[1] 0.01552285 -->
# <!-- # ``` -->
```

### CForest
```{r run model}
require(party)
crf <- cforest(as.factor(effect_10) ~ size.length.um.used.for.conversions + particle.volume.um3 + exposure.duration.d +
                 media.temp + dose.particles.mL.master,
               controls = cforest_control(ntree = 500,
                                          mincriterion = qnorm(0.8), 
                                          trace = TRUE), # adds project bar because it's very slow
               data = multiVar_sub)

crf
```

```{r fit results 1}
train2 <- multiVar_sub %>% dplyr::select(size.length.um.used.for.conversions,
                                  particle.volume.um3,exposure.duration.d,media.temp,dose.particles.mL.master)
train3 <-multiVar_sub %>% dplyr::select(size.length.um.used.for.conversions,
                                  particle.volume.um3,exposure.duration.d,media.temp,dose.particles.mL.master,
                                  effect_10)

fitted <- predict(crf, train2, OOB = TRUE, type ="response")
#rpart.prob <- predict(t1, newdata=imputedSmalls.requested.voluntary,type="prob")

misClasificError <- mean(fitted != train3$effect_10)
print(paste('Training Accuracy', 1 - misClasificError))
```

```{r}
print(crf)
```

<!-- # ```{r} -->
<!-- # #Obtaining predicted probabilites for Test data -->
<!-- # tree.probs=predict(crf, -->
<!-- #                  newdata = train2, -->
<!-- #                  type="prob") -->
<!-- #  -->
<!-- # #Calculate ROC curve -->
<!-- # rocCurve.tree <- roc(train3$effect_10,tree.probs[,'1']) -->
<!-- # #plot the ROC curve -->
<!-- # plot(rocCurve.tree,col=c(4)) -->
<!-- # ``` -->

Alternative ROC Curve
```{r}
require(pROC)
predicted <- predict(crf, train2, OOB=TRUE, type= "response")
auc(as.numeric(train3$effect_10), as.numeric(predicted))
#Calculate ROC curve
rocCurve.tree <- roc(train3$effect_10,as.numeric(predicted))
#plot the ROC curve
plot(rocCurve.tree,col=c(4))
```

```{r}
# compute in-sample results
caret::confusionMatrix(fitted,as.factor(train3$effect_10))
```

```{r}
#plot feature importance
cforestImpPlot <- function(x) {
  cforest_importance <<- v <- varimp(x)
  dotchart(v[order(v)])
}

importancePlot <- cforestImpPlot(crf)
importancePlot
```

<!-- # ```{r} -->
<!-- # crf.prob <- matrix(unlist(crfsrvy.prob), ncol=2, byrow=TRUE) -->
<!-- # apply(crf.prob,2,mean) -->
<!-- # #[1] 0.2524514 0.7475486 -->
<!-- # tab <- round(cbind(by(rpart.prob[,2], INDICES=t1$where, mean), by(crf.prob[,2], INDICES=t1$where, mean)), 4) -->
<!-- # colnames(tab) <- c("rpart", "cforest") -->
<!-- # kable(tab) -->
<!-- # ``` -->

#GLMM
```{r}
#tutorial https://aosmith.rbind.io/2020/08/20/simulate-binomial-glmm/
require(lme4)

mod = glm(effect_f ~ size.length.um.used.for.conversions + polymer + particle.volume.um3 + density.mg.um.3 + organism.group + bio.org + treatments + effect_f + exposure.duration.d + exposure.route + lvl1_f + dose.mg.L.master, 
            data = aoc_z,
            family = binomial(link = "logit") )
mod
```

```{r}
bin_glmm_fun = function(n_sites = 10,
                        b0 = 0,
                        b1 = 1.735,
                        num_samp = 50,
                        site_var = 0.5) {
     site = rep(LETTERS[1:n_sites], each = 2)
     plot = paste(site, rep(1:2, times = n_sites), sep = "." )
     treatment = rep( c("treatment", "control"), times = n_sites)
     dat = data.frame(site, plot, treatment)

     site_eff = rep( rnorm(n = n_sites, mean = 0, sd = sqrt(site_var) ), each = 2)

     log_odds = with(dat, b0 + b1*(treatment == "treatment") + site_eff)
     prop = plogis(log_odds)
     dat$num_samp = num_samp
     dat$y = rbinom(n = n_sites*2, size = num_samp, prob = prop)

     glmer(cbind(y, num_samp - y) ~ treatment + (1|site),
           data = dat,
           family = binomial(link = "logit") )
}

set.seed(16)
bin_glmm_fun()

# Random Forest Modelling

## Preparation
```{r}
library(quantregForest)
library(caret)
library(tidyverse)
library(tidymodels)
library(skimr)
library(sf)
library(ggspatial)
library(nhdplusTools)
library(patchwork)
library(Metrics)
library(gt)
```

```{r}
aoc_z_2 <- aoc_z %>% 
  mutate_if(is.character, as.factor)# %>%
  # mutate(effect_10 = case_when( #convert ordinal to numeric
  #    effect_f == "Yes" ~ 1,
  #    effect_f == "No" ~ 0))

#choose relevant predictors and log-transform
multiVar_small_log <- aoc_z_2 %>% 
  dplyr::select(size_f,
                ID, #row number ID for split/joining by study
                doi, #need to split studies
                size.length.um.used.for.conversions, shape, polymer, particle.volume.um3, density.mg.um.3, organism.group, bio.org, treatments, effect_f, exposure.duration.d, exposure.route, lvl1_f, dose.mg.L.master) %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
  drop_na() %>%  #drop missing
 mutate(effect_10 = case_when( #convert ordinal to numeric
     effect_f == "Y" ~ 1,
     effect_f == "N" ~ 0
   ))# %>% 
 #  select(-effect_f)
skim(multiVar_small_log)
```

```{r}
#no log transform for comparison
multiVar_small <- aoc_z %>% 
  dplyr::select(size_f, size.length.um.used.for.conversions, shape, polymer, particle.volume.um3, density.mg.um.3, organism.group, bio.org, treatments, effect_f, life.stage,
                exposure.duration.d, lvl1_f, dose.mg.L.master, dose.particles.mL.master, dose.um3.mL.master)# %>% 
  #drop_na() #%>%  #drop missing
# mutate(effect_10 = case_when( #convert ordinal to numeric
#     effect == "Y" ~ 1,
#     effect == "N" ~ 0
#   )) %>% 
  # select(-effect)

#ensure completeness
skim(multiVar_small)
```

##NON-LOG TRANSFORMED
### Missing Value Imputation by Rough Fix
```{r}
#rough fix
multiVar_small_roughfix <- na.roughfix(multiVar_small)

# Create calibration and validation splits with tidymodels initial_split() function.
set.seed(4)
multiVar_small_split <- multiVar_small_roughfix %>%
  initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.
# default is 3/4ths split (but 75% training, 25% testing).
# Stratification (strata) = grouping training/testing sets by region, state, etc.
# Using the "strata" call ensures the number of data points in the training data is equivalent to the proportions in the original data set. (Strata below 10% of the total are pooled together.)

# Create a training data set with the training() function
# Pulls from training and testing sets created by initial_split()
multiVar_small_train <- training(multiVar_small_split)
multiVar_small_test <- testing(multiVar_small_split)
# Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.

# Create a separate dataset of available IDs that were not used in the training dataset.
notTrain <- aoc %>% # all COMIDS from StreamCat data, sampled or not
  filter(!ID %in% aoc$ID) # Removing sites used to train the model. n = 140,097

count_optimized <- paste0('n = ',nrow(multiVar_small_roughfix))

skim(multiVar_small_roughfix)
```
#### Kitchen Sink Model
#### LINEAR DATA
```{r}
# Step Three - Kitchen Sink model -----------------------------------------
# Random forest -- 
# a decision tree model, using predictors to answer dichotomous questions to create nested splits.
# no pruning happens - rather, multiple trees are built (the forest) and then you are looking for consensus across trees
# training data goes down the tree and ends up in a terminal node.
# if testing data goes down the same route, then this upholds our conclusions. Or, if it goes awry, this allows us to look for patterns in how it goes awry.

set.seed(2) # assures the data pulled is random, but sets it for the run below (makes outcome stable)
myrf <- randomForest(y = multiVar_small_roughfix$effect_f, # dependent variable
  x = multiVar_small_roughfix %>%
    dplyr::select(-effect_f), # selecting all predictor variables
  importance = T, # how useful is a predictor in predicting values (nothing causal)
  proximity = T, 
  na.action = na.roughfix,
  ntrees = 100) # 500 trees default.

myrf # examine the results.
```
~20% error rate. Let's compare this with the same model with log-transformed values.

##### Predictor Selector/Recursive Feature Elimination
```{r}
# Using caret to select the best predictors
# What are the parameters you want to use to run recursive feature elimination (rfe)?
my_ctrl <- rfeControl(functions = rfFuncs,
                      method = "cv",
                      verbose = FALSE,
                      returnResamp = "all")

# rfe = recursive feature elimination
# THIS STEP TAKES FOR-EV-ER!!!
set.seed(22)
my_rfe <- rfe(y = multiVar_small_roughfix$effect_f, # set dependent variable
              x = multiVar_small_roughfix %>% 
                dplyr::select(-effect_f),
              rfeControl = my_ctrl,
               size = c(1:2, 4, 6, 8, 10, 12, 14, 15))#,
             # na.action = na.roughfix())

# sets how many variables are in the overall model
#               # I have 13 total possible variables, so I've chosen increments of 3 to look at.
#               rfeControl = my_ctrl,
#               na.action = na.roughfix,
#               testX = multiVar_small_test %>% dplyr::select(-effect_f),
#               testY = multiVar_small_test$effect_f)

# can you make your model even simpler?
# the following will pick a model with the smallest number of predictor variables based on the tolerance ("tol") that you specify (how much less than the best are you willing to tolerate?)

#inspect
my_rfe
```

```{r}
trellis.par.set(caretTheme())
#visualize
plot(my_rfe,type = c("g", "o"))
```

```{r}
my_size <- pickSizeTolerance(my_rfe$results, metric = "Accuracy", tol = 5, maximize = TRUE)
# higher tol (~10) gives you less variables
# lower tol (~1) gives you more variables - "I'd like the simplest model within 1% of the best model."
pickVars(my_rfe$variables, size = my_size)
```
The pickSizeTolerance determines the absolute best value then the percent difference of the other points to this value. This approach can produce good results for many of the tree based models, such as random forest, where there is a plateau of good performance for larger subset sizes. For trees, this is usually because unimportant variables are infrequently used in splits and do not significantly affect performance.

>>>>

---
title: "SVM, K-NN"
author: "Scott Coffin"
date: "2/18/2021"
output: 
 html_document:
    code_folding: hide
    theme: sandstone
    toc: yes
    toc_float: yes
    toc_depth: 4
    number_sections: true
  word_document:
    toc: yes
---
```{r global-options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      warning=FALSE, message=FALSE, echo = FALSE)
```
---

# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load packages
library(tidyverse) #General everything
library(RColorBrewer)
library(ggplot2) #General plotting
library(ggrepel) #For adding text labels that repel away from data points
library(calecopal) #Color palette
library(shiny) #Runs shiny
library(shinythemes) #Shiny theme for the page
library(shinyWidgets) #Widgets
library(scales) #SSD - Use the percent format
library(reshape2) #Overview tab - melts bars together
library(ssdtools) #SSD package
library(DT) #Build HTML data tables
#library(plotly) #Make plots interactive
library(viridis) #Colors
library(scales) #To use "percent" function
library(tigerstats) #turns things into percents
library(ggbeeswarm) #plot all points
library(fitdistrplus, exclude = 'select') #alt SSD 
library(RColorBrewer) #colors
library(pheatmap) #pretty heat maps
library(rpart)  #for trees
#library(rattle)    # Fancy tree plot This is a difficult library to install (https://gist.github.com/zhiyzuo/a489ffdcc5da87f28f8589a55aa206dd) 
library(rpart.plot)             # Enhanced tree plots
library(RColorBrewer)       # Color selection for fancy tree plot
library(party)                  # Alternative decision tree algorithm
# library(partykit)               # Convert rpart object to BinaryTree
library(pROC)   #for ROC curves
library(uwot) #umap
library(ISLR)  #for the Carseat Data
library(lme4) #general linearized mixed model GLMM
library(quantregForest)
library(caret) #support vector machines
library(tidyverse)
library(tidymodels)
library(skimr)
library(sf)
library(ggspatial)
library(nhdplusTools)
library(patchwork)
library(Metrics)
library(gt)
library(randomForest)

# Attach Packages
library(kernlab)      # SVM methodology
library(e1071)        # SVM methodology
```
## Data import
```{r data import}
# Load finalized dataset.
aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000) %>% 
  mutate(ID = paste0("ID",row_number()))

@@

multiVar <- aoc_z %>% dplyr::select(#doi, size.category, 
                                    size_f,
                                    size.length.um.used.for.conversions, 
                                    shape, 
                                    polymer, 
                                    particle.volume.um3, 
                                    density.mg.um.3, 
                                    organism.group,
                                    environment, 
                                    bio.org, #biological level of organization
                                    #af.time, #assessment factor based on exposure time
                                    treatments, #number of doses (no including control)
                                    effect, #yes no
                                    effect_f,
                                    effect_10,
                                    size_f,
                                    exposure.duration.d, 
                                    exposure.route, #Factor
                                    organism.group, #factor
                                    media.temp, #numeric
                                    lvl1_f, #endpoints
                                    lvl2_f, #endpoints
                                   # lvl3, 
                                     dose.mg.L.master, 
                                    sex, #factor
                                    media.ph, #numeric
                                    media.sal.ppt, #numeric
                                    dose.particles.mL.master,
                                   effect.metric, #NOEC LOEC
                                    functional.group, #factor
                                    charge, #positive or negatibe
                                    zetapotential.mV, # numeric   
                                   max.size.ingest.mm,#max ingestible size
                                   acute.chronic_f,
                                   dose.mg.L.master.AF.noec,
                                   dose.particles.mL.master.AF.noec,
                                   max.size.ingest.mm, #maximum ingestible size range
                                   effect.score) %>%  #1 = minor, 2 = photosynthesis, feeding, 3 = growth, chlorophyll content, 4 = reproduction, 5  = population growth, 6 = survival
  filter(!size_f == "Not Reported") %>%   #take out not reported 
                                 #  max.size.ingest.mm) %>%  #max ingestible size
  filter(!size_f == "Not Reported")  #take out not reported

@@

# Support Vector Machine
## Kitchen Sink
### Visualization

#### Support Vector Machine - Non-linear

Support Vector Classifiers are a subset of the group of classification structures known as Support Vector Machines. Support Vector Machines can construct classification boundaries that are nonlinear in shape. The options for classification structures using the svm() command from the e1071 package are linear, polynomial, radial, and sigmoid. Toobserve relationships of classifiers, we plot the data in two dimensions.

```{r}
# Plot data
multiVar %>% 
ggplot(aes(x = log(size.length.um.used.for.conversions), y = log(dose.mg.L.master), color = effect_f, shape = effect_f)) + 
  geom_point(size = 2) +
  scale_color_manual(values=c("#000000", "#FF0000")) +
  theme(legend.position = "none")
```
Notice that the data is not linearly separable, and furthermore, isn’t all clustered together in a single group. There are two sections of class 1 observations with a cluster of class 2 observations in between. To demonstrate the power of SVMs, we’ll take 100 random observations from the set and use them to construct our boundary. We set kernel = "radial" based on the shape of our data and plot the results.

```{r}
# set pseudorandom number generator
set.seed(123)
# sample training data
set.seed(4)
kitchenSink_split <- kitchenSink %>%
  initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.
kitchenSink_train <- training(kitchenSink_split)
kitchenSink_test <- testing(kitchenSink_split)
# Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.
count_kitchenSink <- paste0('n = ',nrow(kitchenSink))
skim(kitchenSink)

subTrain <- kitchenSink_test %>% 
  dplyr::select(c(effect_f, dose.mg.L.master, size.length.um.used.for.conversions)) %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10)

#fit model
svmfit_kitchenSink <- svm(effect_f ~ ., 
                          data = subTrain, kernel = "radial", gamma = 1, cost = 1)
summary(svmfit_kitchenSink)
# plot classifier
plot(svmfit_kitchenSink, data = subTrain)
```
The same procedure can be run using the kernlab package, which has far more kernel options than the corresponding function in e1071. In addition to the four choices in e1071, this package allows use of a hyperbolic tangent, Laplacian, Bessel, Spline, String, or ANOVA RBF kernel. To fit this data, we set the cost to be the same as it was before, 1.
```{r}
# Fit radial-based SVM in kernlab
kernfit <- ksvm(effect_f ~ ., data = subTrain, type = "C-svc", kernel = 'rbfdot', C = 1, scaled = c())
# Plot training data
plot(kernfit, data = subTrain)
```

### Data Split
```{r}
#choose relevant predictors and log-transform
kitchenSink <- aoc_z %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate(doi = as.character(doi)) %>% 
  dplyr::select(size_f,
                #doi, #need to split studies
                size.length.um.used.for.conversions, shape, polymer, particle.volume.um3, density.mg.um.3, organism.group, bio.org, treatments, effect_f, exposure.duration.d, exposure.route, lvl1_f, dose.mg.L.master) %>%
  # mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
 # mutate(effect_10 = case_when( #convert ordinal to numeric
 #     effect_f == "Y" ~ 1,
 #     effect_f == "N" ~ 0
 #   ))
drop_na() #drop missing

skim(kitchenSink)
```
#### Drop NA
```{r}
# Create calibration and validation splits with tidymodels initial_split() function.
set.seed(4)
kitchenSink_split <- kitchenSink %>%
  initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.
# default is 3/4ths split (but 75% training, 25% testing).
# Stratification (strata) = grouping training/testing sets by region, state, etc.
# Using the "strata" call ensures the number of data points in the training data is equivalent to the proportions in the original data set. (Strata below 10% of the total are pooled together.)

# Create a training data set with the training() function
# Pulls from training and testing sets created by initial_split()
kitchenSink_train <- training(kitchenSink_split)
kitchenSink_test <- testing(kitchenSink_split)
# Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.

count_kitchenSink <- paste0('n = ',nrow(kitchenSink))

skim(kitchenSink)
```

#### Rough Fix NA
```{r}
#rough fix
kitchenSink_roughfix <- na.roughfix(kitchenSink)

# Create calibration and validation splits with tidymodels initial_split() function.
set.seed(4)
kitchenSink_split_roughfix <- kitchenSink_roughfix %>%
  initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.
# default is 3/4ths split (but 75% training, 25% testing).
# Stratification (strata) = grouping training/testing sets by region, state, etc.
# Using the "strata" call ensures the number of data points in the training data is equivalent to the proportions in the original data set. (Strata below 10% of the total are pooled together.)

# Create a training data set with the training() function
# Pulls from training and testing sets created by initial_split()
kitchenSink_train_roughfix <- training(kitchenSink_split_roughfix)
kitchenSink_test_roughfix <- testing(kitchenSink_split_roughfix)
# Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.

count_roughfix <- paste0('n = ',nrow(kitchenSink_roughfix))

skim(kitchenSink_roughfix)
```

#### Multiple Imputation NA
